\textbf{NBA Party}

### 2017-07-12 cardinal_analytx_presentation/nightlife_and_sports.tex
Updated tables. Moved slides around.

\begin{frame}   
  \frametitle{Why is this problem hard?}
  \begin{block}{Partying is a latent variable.}
    Instead, we proxy for the possibility of athletes partying by looking at 
    play following games in party cities.
  \end{block}
  \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
    We colloquially call cities with high levels of nightlife ``party cities''.   \end{block}
\end{frame}

\begin{frame}   \frametitle{What's the specific question, exactly?}
  \begin{block}{Do teams/athletes perform worse the day after playing in a party-oriented city?\footnote{  We will define ``perform worse'' relative to bookmaker expectations, which are in part driven by
  bettors' expectations.}}
\vspace{12pt}     I.e. we are concerned with ``hangover effects''.
  \end{block}
\end{frame}

@@

\begin{frame}   \frametitle{Identifying variation}
  \begin{block}{Performance as a function of previous day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
    then looking at the \emph{actual outcome} of the game $\leadsto$ identification.

    \vspace{15pt}

    Weaker, still: if odds of beating the spread uncorrelated with previous game location
    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
\end{frame}

@@

    \vspace{12pt}
    Helps to account for systematically running into teams the day after visitation to 
    a party oriented city who are particularly (not)-skilled.
  \end{block}
\end{frame}

@@

  % \begin{itemize}     \item If bookkeepers aware of hangover effects, our variable of interest may not appear significant, in spite of there being a true effect.

  %   \item If bookkeepers systematically mis-value teams which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.

  %     (Given size of betting market, this is unlikely)
  %   \end{itemize}

@@

\vspace{12pt}  Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
  who work harder on the court and expend energy boxing out other players will have higher
  rebound rates. \end{frame}

% \begin{frame}   \frametitle{Analysis: rebounds}
%   \begin{itemize}     \item For each player and each game, we calculate the % number of rebounds obtained per-minute of game-play
%     \item From each observation, we subtract the player's average number of % rebounds per minute taken across the entire season.
%     \item We then run a linear regression explaining demeaned rebounds as a % function of 
%       nightlife visitation and other features describing player fatigue.   % \end{itemize} \end{frame}

% \begin{frame}   \frametitle{Results: rebounds}
%   \centering
%   \tiny{ % m2
% \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & % \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & % Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & % $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 % \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    % & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   % Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations % & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std. % Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  % \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; % $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}

\begin{frame}   \frametitle{Hangover correlates with allowing more points}
  We examine points allowed and points scored (on the team level).

  \vspace{12pt}Common knowledge suggests that overall defense of a team is strongly influenced by individual player effort.

  \vspace{12pt}Is it the case the hangover effects are more prevalent on defensive aspects of gameplay? \end{frame}

\begin{frame}   \frametitle{Defense in the NBA}
  \begin{adjustwidth}{-.75cm}{}
  \tiny{  % mpa1, tpa, tps
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.649$^{***}$ &  &  \\    & (0.718) &  &  \\    & & & \\   Continuous party measure &  & 1.833$^{***}$ & 0.447$^{*}$ \\    &  & (0.246) & (0.245) \\    & & & \\   Lag change in possessions & 1.058$^{*}$ & 0.022$^{***}$ & 0.035$^{***}$ \\    & (0.615) & (0.007) & (0.007) \\    & & & \\   Logged travel distance & $-$0.045 & $-$0.028 & $-$0.213$^{***}$ \\    & (0.038) & (0.074) & (0.074) \\    & & & \\   Number hours since last game & 0.009$^{**}$ & 0.015$^{***}$ & 0.018$^{***}$ \\    & (0.004) & (0.003) & (0.003) \\    & & & \\   Home team effect & $-$2.997$^{***}$ & $-$3.412$^{***}$ & 2.211$^{***}$ \\    & (0.218) & (0.116) & (0.116) \\    & & & \\   Constant & $-$3.429 & 101.941$^{***}$ & 100.112$^{***}$ \\    & (2.926) & (0.479) & (0.477) \\    & & & \\  \hline \\[-1.8ex]  Observations & 14,843 & 44,689 & 44,689 \\  R$^{2}$ & 0.018 & 0.021 & 0.010 \\  Adjusted R$^{2}$ & 0.017 & 0.021 & 0.010 \\  Residual Std. Error & 11.070 (df = 14831) & 12.002 (df = 44683) & 11.967 (df = 44683) \\  F Statistic & 24.021$^{***}$ (df = 11; 14831) & 196.086$^{***}$ (df = 5; 44683) & 92.809$^{***}$ (df = 5; 44683) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
\end{adjustwidth} 
\end{frame}

>>>>

\begin{frame}   
  \frametitle{Why is this problem hard?}
  \begin{block}{Partying is a latent variable.}
    Instead, we proxy for the possibility of athletes partying by looking at 
    play following games in party cities.
  \end{block}
  \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
    We dub cities with high levels of nightlife ``party cities''.  
  \vspace{12pt}
  \begin{block}{Another is how to measure a dip in performance?}
    We need to consider player ability, team momentum, etc.
  \end{block} \end{block}
\end{frame}

\begin{frame}   \frametitle{What's the specific question, exactly?}
  \begin{block}{Do teams/athletes perform worse the day after playing in a party-oriented city?}
\vspace{12pt}     I.e. we are concerned with ``hangover effects''.

\vspace{12pt}

% E.g. a team visits Las Vegas for the day. How is their next game?
  \end{block}
\end{frame}

@@

\begin{frame}   \frametitle{Identifying variation}
  \begin{block}{Performance as a function of previous day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
    then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
  \end{block}
\end{frame}

    Helps to account for systematically running into teams the day after visitation to 
    a party oriented city who are particularly (not)-skilled.
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Implications and pitfalls of using a black-box feature}
  \begin{block}{If bookkeepers aware of hangover effects...}     then our variable of interest may not appear significant, in spite of there being a true effect.   \end{block}
  \vspace{12pt}
  \begin{block}{If bookkeepers systematically mis-value teams...}
     which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.

      (Given size of betting market, this is unlikely)
    \end{block}   
\end{frame}

@@

@@

@@

  \vspace{12pt}  
  \begin{block}{Is it the case the hangover effects are more prevalent on defensive aspects of gameplay?}     Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
    who work harder on the court and expend energy boxing out other players will have higher
    rebound rates.    \end{block}
\end{frame}

\begin{frame}   \frametitle{Defense in the NBA}
  \tiny{  % mpa1, tpa, tps
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & \multicolumn{2}{c}{Team Points Admitted} & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.970$^{***}$ &  &  \\    & (0.835) &  &  \\    & & & \\   Continuous party measure &  & 1.661$^{***}$ & $-$0.139 \\    &  & (0.644) & (0.637) \\    & & & \\   Lag change in possessions & 6.694$^{***}$ & 6.751$^{***}$ & 5.858$^{***}$ \\    & (0.973) & (0.973) & (0.963) \\    & & & \\   Logged travel distance & $-$0.114 & $-$0.132 & $-$0.285 \\    & (0.198) & (0.198) & (0.196) \\    & & & \\   Number hours since last game & 0.011$^{*}$ & 0.015$^{**}$ & 0.016$^{**}$ \\    & (0.006) & (0.007) & (0.007) \\    & & & \\   Home team effect & $-$3.269$^{***}$ & $-$3.269$^{***}$ & 2.417$^{***}$ \\    & (0.303) & (0.303) & (0.300) \\    & & & \\   Constant & 68.114$^{***}$ & 67.686$^{***}$ & 71.348$^{***}$ \\    & (4.805) & (4.810) & (4.758) \\    & & & \\  \hline \\[-1.8ex]  Observations & 6,234 & 6,234 & 6,234 \\  R$^{2}$ & 0.091 & 0.090 & 0.084 \\  Adjusted R$^{2}$ & 0.087 & 0.086 & 0.079 \\  Residual Std. Error (df = 6201) & 11.670 & 11.675 & 11.549 \\  F Statistic (df = 32; 6201) & 19.453$^{***}$ & 19.247$^{***}$ & 17.720$^{***}$ \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
\end{frame}

\begin{frame}
  \centering
  \includegraphics[scale=0.25]{../writing/distribution_point_spreads_nba}
\end{frame}
